Best Practices for Alchemical Free Energy Calculations: v\versionnumber

### 2018-07-30 paper/manuscript.tex
Add short section about the structural definition of the bound state and weak binders.

@@

\subsubsection{Absolute free energy calculations must handle the standard state and use restraints}

\todo[inline, color={red!40}]{AR, LNN: complete absolute free energy simulation protocol}

@@

Absolute free energy calculations are typically done with respect to a specific reference state or reference concentration.
For solvation free energy calculations this is typically straightforward to handle and treating it correctly simply means ensuring that the non-interacting solute is taken to the same (or equivalent) final reference state in both environments, e.g. that the transformation involves a 1 M to 1M equivalent transfer free energy (where the non-interacting solute still occupies essentially the same volume as the solute in the interacting system).
So typically in such cases no special care is required to ensure the correct standard state, as long as the \emph{experimental} data being analyzed uses the same standard state and if it does not, a simple entropic correction is needed.

However, for binding the situation is much more complex and requires special care.
Experimental absolute binding free energies are reported relative to a specific reference state -- a 1 M standard state -- which must also be used in calculations.
In practice this has implications for how the calculations are done, as the reference concentration must enter the thermodynamic cycle employed.

>>>>

\paragraph{Structural definition of the bound state and weak binders.}
\todo[inline, color={red!40}]{AR: This paragraph presents concepts that are common to absolute and relative binding free energy calculations, but it is useful to have this \textit{before} talking about restraints. Should we move the "common" section before the "differences" section?}

In binding free energy calculations, extra care should be taken when simulating the bound state, especially when dealing with weak binders and absolute free energy calculations in combination with enhanced sampling techniques.
In principle, only configurations of the receptor-ligand complex that we consider "bound" should be sampled from the bound state, and, in atomistic simulations, this requires to establish a structural definition of the bound state.
In practice, the simulation of the bound state starts with the ligand already placed in the binding site and relies on kinetic trapping to maintain a bound complex.
However, this strategy may not be sufficient if the complex dissociation rate is high or if methodologies such as Hamiltonian replica exchange [ref] and expanded ensemble [ref] are employed in absolute free energy calculations to enhance sampling since, in both cases, the ligand may find a way out of the binding site on timescales that are achievable by modern MD simulations.
A solution commonly adopted in these cases is the use of one or more restraints making the unbound configurations energetically unfavorable through the addition of extra terms in the potential function.
It should be noted that these additional energy terms can introduce a bias in the predicted free energy that should be properly handled (see Section \ref{sec:standardstate-restraints} for details)\todo[inline, color={red!40}]{AR: Maybe point to the specific section that talks about the restraint protocol and how to remove the bias.}.
In practice, even with restraints, it is not always trivial to force a molecular dynamics simulation to explore a restricted region of the configurational space with a complicated geometry such as a binding site.
If the ligand is a tight binder\todo[inline, color={red!40}]{AR: Should we give an order of magnitude to define a tight and a weak binder?}, it is usually safe to employ a restraint that allows some of the unbound configurations to be sampled since they generally do not contribute significantly to the partition function (i.e., they have a relatively small Boltzmann weight) as long as the sampled volume is not so large that their cumulative contribution becomes significant.
However, this is more problematic when dealing with weak binders as the unbound configurations can have a non-negligible Boltzmann weight, and their binding affinity can exhibit a significant dependency on the restraint type and parameters that are used to determined the sampled volume.

Finally, it is useful to keep in mind that, for a meaningful comparison between computational predictions and experimental measurements, the definition of the bound state should be consistent.
In particular, this means that the signal used by the experimental methodology to determine the fraction of bound complexes in solution should in principle reflect the population of complexes in the bound state as defined in the calculation.

@@

\subsubsection{Absolute free energy calculations must handle the standard state and use restraints}
\label{sec:standardstate-restraints}

\todo[inline, color={red!40}]{LNN: complete discussion about Boresch restraints}

\todo[inline, color={red!40}]{AR: would the explanation in the next paragraph be better suited for a more general section?}

@@

Absolute free energies are typically reported with respect to a specific reference or standard state, which effectively determines the arbitrary point at which the free energy is 0.
The role of the standard state is particularly evident with binding free energies, in which having a reference state allows us to obtain a well-defined partial molar free energy for the reaction
\begin{equation*}
\ce{AB <=> A + B}
\end{equation*}
through the well-known expression derived from the law of mass action
\begin{equation} \label{eq:DGfromKAB}
\Delta G = -RT ~ \text{ln} \left( C^0 K_{AB} \right)  = -RT ~ \text{ln}\left( \frac{C^0 C_{AB}}{C_A C_B} \right) ,
\end{equation}
where $R$ is the gas, $T$ is the temperature, $C_X$ is the equilibrium concentration of the chemical species $X$ in the reaction solvent, and the reference state concentration $C^0$ converts the binding constant $K_{AB}$ into a dimensionless quantity expressed in reference concentration units.
It should be noted that ignoring the term $C^0$ is equivalent to assuming a reference concentration of 1~D$^{-1}$, where D are the units used to express $K_{AB}$, and would thus cause the value of $\Delta G$ to vary with the choice of the units.
Typically, it is convenient to define a standard state at a constant pressure of 1~atm and where each chemical species (i.e., A, B, and AB) in the reaction solvent has a concentration of $C^0$~=~1~M~=~1~molecule/1660~\r{A}$^3$ but do not interact with other molecules of A, B, or AB.

\paragraph{Handling the standard state in absolute free energy calculations.}

For solvation free energy calculations, handling the standard state is typically straightforward, and treating it correctly simply means ensuring that the non-interacting solute is taken to the same (or equivalent) final reference state in both environments, e.g. that the transformation involves a 1M to 1M equivalent transfer free energy (where the non-interacting solute still occupies essentially the same volume as the solute in the interacting system).
So typically in such cases no special care is required to ensure the correct standard state, as long as the \emph{experimental} data being analyzed uses the same standard state and if it does not, a simple entropic correction is needed.

However, for binding the situation is much more complex and requires special care.
Experimental absolute binding free energies are reported relative to a specific reference state -- a 1 M standard state -- which must also be used in calculations.
In practice, this has implications for how the calculations are done, as the reference concentration must enter the thermodynamic cycle employed.
